---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
setwd("c:/Users/Ola/Desktop/Coursera/Reproductible_research/Project1")
raw<-read.csv("activity.csv", header=T, na.strings="NA")
raw$Date <- as.Date(raw$date, format="%Y-%m-%d")

```

## What is mean total number of steps taken per day?
```{r readPackage, message=FALSE}
#Reading the dplyr and lattice package 
library(dplyr)
library(lattice)
```

```{r convertingData,fig.height=3}
#and converting the dataset to dplyr dataframe.
data<-tbl_df(raw)

#Calculating the table summarizing the data by day
groupByDay<-group_by(data,Date)
summaryByDate<-summarize(groupByDay,StepsPerDay=sum(steps))

#Plotting the histogram
with(summaryByDate, hist(StepsPerDay, main="Histogram of steps taken per day",
                         col="blue", xlab="Steps taken per day"))
```
``` {r calculatingMeanAndMedianQuestion1}
#Calculating the mean and the median
meanNumberOfSteps<-round(mean(summaryByDate$StepsPerDay, na.rm=T),2)
medianNumberOfSteps<-round(median(summaryByDate$StepsPerDay, na.rm=T),2)

```

**Mean** number of steps taken per day is `r meanNumberOfSteps`. The **median** is almost the same and equals `r medianNumberOfSteps`.

```{r removingUnnecessaryQuestion1, echo=FALSE}
rm(groupByDay, summaryByDate)
```



## What is the average daily activity pattern?
``` {r summarizingDataQuestion2}
#Calculating the table summarizing the data by day
dataWoNa<-filter(data, is.na(steps)==FALSE)
groupByInterval<-group_by(dataWoNa,interval)
summaryByInterval<-summarize(groupByInterval,MeanIntervalByDay=mean(steps))
```
```{r plottingQuestion2, fig.height=3}
with(summaryByInterval, plot(interval, MeanIntervalByDay, type="l", main="Average daily activity pattern", xlab="Intervals", ylab="Average steps per day"))

maxNumberOfSteps<-max(summaryByInterval$MeanIntervalByDay)
intervalWithMaxSteps<-as.integer(filter(summaryByInterval,MeanIntervalByDay==maxNumberOfSteps)[1])
maxAvgSteps<-as.integer(filter(summaryByInterval,MeanIntervalByDay==maxNumberOfSteps)[2])

abline(v=intervalWithMaxSteps, col="blue", lty=3)
abline(h=maxAvgSteps, col="blue", lty=3)
text(1200,180, paste("Max=" , maxAvgSteps), col="blue")
```

The interval which on average across all the days in the dataset, contains the maximum number of step is `r intervalWithMaxSteps`.

```{r removeUnnecessary2, echo=FALSE}
rm(dataWoNa, groupByInterval,summaryByInterval,maxNumberOfSteps,  intervalWithMaxSteps, maxAvgSteps)
```


## Imputing missing values
```{r calculatingNA}
dataNa<-filter(data, is.na(steps)==TRUE)
naNumber<-as.integer(summarize(dataNa,  count = n())[1])
```
Number of missing values equals `r naNumber`.
```{r inputtingMissingValues}
#To input missing values a mean from the given interval was used.

groupByInterval<-group_by(data,interval)
summaryByInterval<-summarize(groupByInterval,intervalAverage=mean(steps,na.rm=T))
dataAll <- merge(data, summaryByInterval, by.x="interval", by.y="interval")
dataAll<-mutate(dataAll, stepsInputted=ifelse(is.na(steps)==T,intervalAverage,steps))

```
```{r summarizingDataQuestion3, fig.height=3}
#converting the dataset to dplyr dataframe.
dataAll<-tbl_df(dataAll)

#Calculating the table summarizing the data by day
groupByDay<-group_by(dataAll,Date)
summaryByDate<-summarize(groupByDay,inputtedStepsPerDay=sum(stepsInputted))

#Plotting the histogram
with(summaryByDate, hist(inputtedStepsPerDay, main="Histogram of steps taken per day",
                         col="blue", xlab="Steps taken per day"))
```

``` {r calculatingMeanAndMedianQuestion3}
#Calculating the mean and the median
meanInputtedSteps<-round(mean(summaryByDate$inputtedStepsPerDay),2)
medianInputtedSteps<-round(median(summaryByDate$inputtedStepsPerDay),2)

```
**Mean** number of steps taken per day is `r meanInputtedSteps`. The **median** is the same and equals `r medianInputtedSteps`.

```{r removingQuestion3}
rm(groupByDay, groupByInterval, summaryByDate, summaryByInterval)
```

## Are there differences in activity patterns between weekdays and weekends?
``` {r question4}
Sys.setlocale("LC_ALL","C")
dataAll$Weekdays<-weekdays(dataAll$Date)
dataAll<-mutate(dataAll, Weekends=ifelse(Weekdays=="Saturday"|Weekdays=="Sunday","weekend","weekday"))

#summarizing data for the chart
groupedByInterval<-group_by(dataAll, interval, Weekends)
summaryByInterval<-summarize(groupedByInterval, averageInputtedSteps=mean(stepsInputted))
```
``` {r chartQuestion4, message=FALSE}
chart<-xyplot(averageInputtedSteps ~ interval | Weekends, data=summaryByInterval, 
              layout=c(1,2), type="l", xlab="Interval", ylab="Number of steps",
                  )
print(chart) 
 ```
